---
title: "Intensifiers: Study 1a"
author: "Erin Bennett"
output: 
  html_document:
      toc: false
---

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=F, warning=F, cache=F, message=F, sanitiz =T, fig.width = 5, fig.height = 3)
```

```{r load_settings}
source("~/Settings/startup.R")
```

```{r}
## load intensifier data
unigrams = read.csv("web_1grams.csv")
total_ngrams = 1024908267229
freq = unigrams$frequency
syll = unigrams$syllables
names(freq) = names(syll) = unigrams$ngram

novel_intensifiers = c(
  "lopusly", "fepolopusly",
  "ratumly", "gaburatumly",
  "bugornly", "tupabugornly"
)

## load experiment data
raw.df = read.csv("../../../data/study3_data.csv")
full_df = raw.df %>%
  filter(language %in% c(
    "English",  "english",  "Engli",    "Englsih",  "engish",   "English ",
    "ENGLISH", "ENG", "eng") &
      asses != "No") %>%
  mutate(price = response,
         intensifier = as.character(adverb)) %>%
  select(workerid, price, intensifier, object) %>%
  mutate(logprice = as.numeric(log(price))) %>%
  group_by(object, workerid) %>%
  mutate(logprice = as.numeric(log(price)),
         logprice.centered = logprice - mean(logprice),
         z = as.numeric(scale(log(price))),
         z = ifelse(is.nan(z), 0, z),
         logprice.scaled = z)
df = full_df %>%
  filter(!(intensifier %in% novel_intensifiers)) %>%
  ungroup %>%
  mutate(freq = freq[as.character(intensifier)],
         syll = syll[as.character(intensifier)],
         syll.centered = syll - mean(syll),
         surprisal = -log(freq/total_ngrams),
         surprisal.centered = surprisal - mean(surprisal),
         syll.scaled = scale(syll),
         surprisal.scaled = scale(surprisal),
         word.cost = 0.18124*syll.centered + 0.10704*surprisal.centered)
```

```{r}
lookup_root_of_novel_intensifier = function(word) {
  if (word=="lopusly" | word=="fepolopusly") {
    return("lopus")
  } else if (word=="ratumly" | word=="gaburatumly") {
    return("ratum")
  } else {
    return("bugorn")
  }
}
novel_df = full_df %>% filter(intensifier %in% novel_intensifiers) %>%
  mutate(root = lookup_root_of_novel_intensifier(intensifier),
         length = ifelse(nchar(char(intensifier)) < 9, "short", "long"))
novel_df %>%
  group_by(intensifier, workerid) %>%
  summarise(
    root = root[[1]],
    length = length[[1]],
    logprice.scaled = mean(logprice.scaled)) %>%
  summarise(
    root = root[[1]],
    length = length[[1]],
    low = ci.low(logprice.scaled),
    high = ci.high(logprice.scaled),
    logprice.scaled = mean(logprice.scaled)) %>%
  ggplot(., aes(x=length, y=logprice.scaled)) +
  geom_bar(stat="identity") +
  facet_wrap(~root) +
  geom_errorbar(aes(ymin=low, ymax=high), width=0) +
  xlab("length of nonce word") +
  ylab("normalized log price") +
  ggtitle("Study 3")
ggsave("../edited_draft/images/plot_study3_A.pdf", width=6, height=3)
```



Model:

```{r, echo=T}
m = lmer(logprice.scaled ~ 1 + syll.centered + surprisal.centered +
    (0 + syll.centered + surprisal.centered | workerid) +
    (1 | intensifier), df)
```

```{r}
report = function(m, regressor) {
  coefs = summary(m)$coefficients[regressor,]
  b = coefs[["Estimate"]]
  df = coefs[["df"]]
  t = coefs[["t value"]]
  p = coefs[["Pr(>|t|)"]]
  return(paste(
    "$b=",
    round(b, 3),
    ", t(",
    # "df=",
    round(df, 1),
    ")=",
    round(t, 3),
    ", ",
    ifelse(p<0.05, "p<0.05", paste("p=", round(p, 3), sep="")),
    "$",
    sep=""))
}
```

syllables: `r report(m, "syll.centered")`

surprisal: `r report(m, "surprisal.centered")`

```{r}
novel_m = novel_df %>%
  mutate(length = ifelse(length=="short", -1, 1)) %>%
  lmer(logprice.scaled ~ length +
         (0 + length | workerid) +
         (1 | root), data=.)
summary(novel_m)
```

```{r}
novel_m2 = novel_df %>%
  mutate(length = ifelse(length=="short", -1, 1)) %>%
  lmer(logprice.scaled ~ length + root +
         (0 + length | workerid), data=.)
summary(novel_m2)
anova(novel_m2)
```

```{r}
require(lmtest)
m_simple = lmer(logprice.scaled ~ 1 +
    (0 | workerid) +
    (1 | intensifier), df)
m_syll = lmer(logprice.scaled ~ 1 + syll.centered +
    (0 + syll.centered | workerid) +
    (1 | intensifier), df)
m_surp = lmer(logprice.scaled ~ 1 + surprisal.centered +
    (0 + surprisal.centered | workerid) +
    (1 | intensifier), df)
lr_full = lrtest (m_simple, m)
lr_diff_due_to_surp = lrtest (m_syll, m)
lr_diff_due_to_syll = lrtest (m_surp, m)
```

```{r}
report_chisq = function(df, chisq, p) {
  return(paste(
    "$\\chi^2(",
    # "df=",
    df,
    ")=",
    round(chisq, 3),
    ", ",
    ifelse(p<0.05, "p<0.05", paste("p=", round(p, 3), sep="")),
    "$",
    sep=""))
}
```


full likeihood ratio test: `r report_chisq(lr_full$Df[[2]], lr_full$Chisq[[2]], lr_full[["Pr(>Chisq)"]][[2]])`

likeihood ratio test for surprisal effect beyond syllables effect: `r report_chisq(lr_diff_due_to_surp$Df[[2]], lr_diff_due_to_surp$Chisq[[2]], lr_diff_due_to_surp[["Pr(>Chisq)"]][[2]])`

likeihood ratio test for syllables effect beyond surprisal effect: `r report_chisq(lr_diff_due_to_syll$Df[[2]], lr_diff_due_to_syll$Chisq[[2]], lr_diff_due_to_syll[["Pr(>Chisq)"]][[2]])`

Results:

```{r, width=10}
library(ggrepel)
df %>%
  group_by(intensifier) %>%
  summarise(surprisal.centered = surprisal.centered[[1]],
            syllables = syll[[1]],
            low = ci.low(logprice.scaled),
            high = ci.high(logprice.scaled),
            logprice.scaled = mean(logprice.scaled)) %>%
  ggplot(., aes(x=surprisal.centered,
                y=logprice.scaled,
                colour=syllables)) +
  geom_smooth(method="lm", colour="gray", alpha=0.1) +
  geom_point() +
  geom_text_repel(size=3, aes(label=intensifier)) +
  geom_errorbar(aes(ymin=low, ymax=high), width=0) +
  # geom_text_repel(aes(label=intensifier)) +
  # facet_wrap(~object, scale="free") +
  ylab("normalized log price") +
  scale_colour_gradient(low="gray", high="black") +
  xlab("centered surprisal") +
  ggtitle("Study 3") +
  theme_bw()
# ggsave("../edited_draft/images/plot_study1b.pdf", width=8, height=2.5)
ggsave("../edited_draft/images/plot_study3_B.pdf", width=5, height=3)
```

```{r}
df3 = full_df %>% group_by(intensifier, object) %>%
  summarise(logprice=mean(logprice)) %>%
  summarise(logprice=mean(logprice))
# df1b$intensifier[order(df1b$logprice)] %>%
df3 %>% write.csv("intensifiers_mean_logprice_study3.csv",
                   row.names=F)
```

```{r}
intensities = full_df %>% group_by(intensifier) %>%
  summarise(
    low = ci.low(logprice.scaled),
    high = ci.high(logprice.scaled),
    mean_logprice = mean(logprice.scaled)) %>%
  mutate(
    low = (low - mean(mean_logprice))/sd(mean_logprice),
    high = (high - mean(mean_logprice))/sd(mean_logprice),
    intensity = (mean_logprice - mean(mean_logprice))/sd(mean_logprice)) %>%
  select(intensifier, intensity, low, high)
intensities = intensities[order(intensities$intensity),]
write.csv(intensities, "intensities_study3.csv", row.names=F)
```