---
title: "truly very (non)replication"
author: "erin"
date: "01/26/2015"
output: html_document
---

```{r echo=F}
### load required packages
library(ggplot2, quietly=T)
library(plyr, quietly=T)
library(rjson, quietly=T)
source("~/opt/r_helper_scripts/bootsSummary.r")

### parameters to vary in the analysis
exclude_impatient = T
exclude_manipulation_check = F

d = read.table("~/everything-cocolab/degree-adverbs/experiments/exp11/analysis/exp11.csv", sep=",", header=T)
```

This experiment was run between `r d$assignmentsubmittime[1]` and `r d$assignmentsubmittime[length(d$assignmentsubmittime)]`.

```{r echo=F, fig.width=16, fig.height=16}

# #by worker frequency responses: frequency only
# p = ggplot(d[d$qtype == "frequency" & !is.na(d$qtype),], aes(x=word_type, y=response, fill=word_type)) +
#   geom_bar(stat="identity", position="identity") +
#   geom_text(label=condition)
#   facet_wrap(~ workerid, scale="free") +
#   ylab("frequency") +
#   ggtitle("by worker frequency responses")
# print(p)

# #by worker frequency responses: easier coloring to read, with both frequency and price
d$wt = as.character(d$word_type)
d$wt[d$wt %in% c("bare", "yep")] = "other"
d$wt = as.factor(d$wt)
dodge=position_dodge(.9)
d$lalala = paste( d$qtype, d$word_type)
p = ggplot(d[d$qtype %in% c("frequency", "price") & !is.na(d$qtype),], aes(x=qtype, y=response, fill=wt)) +
  geom_bar(stat="identity", position=dodge) +
 # geom_text(label=condition) +
  facet_wrap(~ workerid, scale="free") +
  ylab("frequency") +
  ggtitle("by worker frequency responses")
print(p)

d = ddply(d, .(workerid), .fun=function(subd) {
  subject_is_impatient = sum(as.logical(subd$impatient[is.na(subd$qtype)]))>1
  subd$impatient = rep(subject_is_impatient, nrow(subd))
  return(subd)
})
d = ddply(d, .(workerid), .fun=function(subd) {
  target_freq = as.numeric(as.character(subd$response[subd$qtype == "frequency" & subd$word_type == "target" & !is.na(subd$response)]))
  control_freq = as.numeric(as.character(subd$response[subd$qtype == "frequency" & subd$word_type == "control" & !is.na(subd$response)]))
  subject_got_manipulation = target_freq > control_freq
  subd$got_manipulation = rep(subject_got_manipulation, nrow(subd))
  return(subd)
})
n_weird_freq_workers = length(unique(d$workerid[!d$got_manipulation]))
n_impatient_workers = length(unique(d$workerid[d$impatient]))
n_failed_both_workers = length(unique(d$workerid[d$impatient & !d$got_manipulation]))
if (exclude_impatient) {
  d = d[d$impatient == F,]
}
if (exclude_manipulation_check) {
  n_total_workers = length(unique(d$workerid))
  d = d[d$got_manipulation == T,]
  n_good_workers = length(unique(d$workerid))
  n_excluded_workers_manipulation = n_total_workers - n_good_workers
}
d$condition = d$Answer.target
d = d[d$qtype %in% c("frequency", "price"),c("workerid", "word_type", "word", "response", "qtype", "condition")]
d$response = as.numeric(as.character(d$response))
d = ddply(d, .(workerid, qtype), .fun=function(subd) {
  subd$diff = subd$response - subd$response[!(subd$word_type %in% c("control", "target"))]
  return(subd)
})
#d$diff[d$qtype == "frequency"] = NA
prices = d[d$qtype == "price",]
freqs = d[d$qtype == "frequency",]

if (exclude_impatient) {
  conj1 = ""
} else {
  conj1 = "not"
}
if (exclude_manipulation_check) {
  conj2 = ""
} else {
  conj2 = "not"
}
```

`r n_impatient_workers` participants clicked continue at least two times within 10s of seeing a comic page. These participants were `r conj1` excluded.

`r n_weird_freq_workers` participants thought the target (repeated) intensifier was less common than the control one. These participants were `r conj2` excluded.

`r n_failed_both_workers` participants failed both checks.

```{r echo=F}
```