---
title: "Power Analysis"
author: "Erin Bennett"
output: 
  html_document:
      toc: false
---

```{r global_options, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo=F, warning=F, cache=F, message=F, sanitiz =T, fig.width = 5, fig.height = 3)
```

```{r load_settings}
source("~/Settings/mini-startup.R")
```

## Experiment

Rating experiment.

## Planned analysis

### 1. Reformat data

We first excluded any participants who are not native speakers of English or who stated that they did not think they followed instructions.

People represent prices logarithmically, so we log-transform all of the prices that participants entered.

We don't care about modeling the prices of the individual objects. We think that this factor might interact with participants (different participants have different beliefs about price distributions for different objects), but we do not have enough data to fit an accurate mixed model with interacting random effects. So we z-score within each participant's ratings for each object (Thanks, Reviewer, for the suggestion!).

```{r}
## load intensifier data
unigrams = read.csv("web_1grams.csv")
total_ngrams = 1024908267229
freq = unigrams$frequency
syll = unigrams$syllables
names(freq) = names(syll) = unigrams$ngram

## load experiment data
df = read.csv("replication2016-data.csv") %>%
  filter(language %in% c("English", "ENGLISH", "ENG", "eng", "english") &
           assess == "Yes") %>%
  select(workerid, price, intensifier, object) %>%
  mutate(workerid = factor(workerid),
         logprice = as.numeric(log(price))) %>%
  group_by(object) %>%
  mutate(logprice.scaled.by.object = as.numeric(scale(logprice)),
         logprice.scaled.by.object = ifelse(
           is.nan(logprice.scaled.by.object), 0,
           logprice.scaled.by.object)) %>%
  ungroup %>%
  group_by(object, workerid) %>%
  mutate(logprice = as.numeric(log(price)),
         logprice.centered = logprice - mean(logprice),
         z = as.numeric(scale(log(price))),
         z = ifelse(is.nan(z), 0, z),
         logprice.scaled = z) %>%
  ungroup %>%
  mutate(freq = freq[as.character(intensifier)],
         syll = syll[as.character(intensifier)],
         syll.centered = syll - mean(syll),
         surprisal = -log(freq),
         surprisal.centered = surprisal - mean(surprisal),
         syll.scaled = scale(syll),
         surprisal.scaled = scale(surprisal),
         word.cost = 0.18124*syll.centered + 0.10704*surprisal.centered)
```

Here's the original data. The slopes and intercepts vary by both participants and objects. In particular, watches seem to have larger slopes than laptops.

```{r}
df %>% ggplot(., aes(x=word.cost,
                     y=logprice,
                     group=paste(workerid, object))) +
  geom_line(stat="smooth",method = "lm", alpha = 0.1) +
  facet_wrap(~object) +
  geom_point(alpha=0.1)
```

And here's what the data look like when we z-score within each participant's responses for each object. Now each participant has an intercept of zero (kind of obviously...), but the slopes vary across participants.

```{r}
df %>% ggplot(., aes(x=word.cost,
                     y=logprice.scaled,
                     group=paste(workerid, object))) +
  geom_line(stat="smooth",method = "lm", alpha = 0.1) +
  facet_wrap(~object) +
  geom_point(alpha=0.1)
```

#### Preprocessing for replication design

We will have a larger set of intensifiers in the replication experiment, so ideally we would like to run a "replication design" where we run a certain number of participants on each of 3 or so disjoint sets of intensifiers.

How would this interact with the zscoring step in preprocessing?

```{r}
shuffled_intensifiers = sample(unique(df$intensifier), replace=F)
shuffled_workers = sample(unique(df$workerid), replace=F)

create_data_subset = function(data, intensifiers, workers) {
  indices = c(apply(
    expand.grid(intensifiers, workers),
    1,
    function(lst) {
      indices = which(df$intensifier==lst[1] &
                        df$workerid==lst[2])
      return(indices)
    }))
  return(data[indices,])
}

create_replication_set = function(set_index) {
  ## split full intensifiers list into 3 disjoint sets
  ## split workers into corresponding 3 disjoint sets
  if (set_index==1) {
    intensifiers_subset = shuffled_intensifiers[1:13]
    workers_subset = shuffled_workers[1:9]
  } else if (set_index==2) {
    intensifiers_subset = shuffled_intensifiers[14:26]
    workers_subset = shuffled_workers[10:18]
  } else {
    intensifiers_subset = shuffled_intensifiers[27:40]
    workers_subset = shuffled_workers[19:28]
  }
  ## sample the rows of the data frame for each set
  data_subset = create_data_subset(df,
                                   intensifiers_subset,
                                   workers_subset) %>%
    mutate(set = set_index)
  return(data_subset)
}

replication_design_data = do.call(
  rbind,
  lapply(1:3, create_replication_set))
```

Here is a subset of the data where each participant sees items from only one of three disjoint sets of intensifiers.

```{r}
replication_design_data %>%
  ggplot(., aes(x=word.cost,
                y=logprice,
                group=paste(workerid, object))) +
  geom_line(stat="smooth",method = "lm", alpha = 0.1) +
  facet_wrap(~object) +
  geom_point(alpha=0.1)
```

And here is that same data, zscored within participant and object. (We have very little data in each normalized subset in this simulation: we will have much more data in the proposed future replication experiment.)

```{r}
replication_design_data %>%
  group_by(object, workerid) %>%
  mutate(
    z = as.numeric(scale(log(price))),
    z = ifelse(is.nan(z), 0, z),
    logprice.scaled = z
  ) %>%
  ungroup %>%
  ggplot(., aes(x=word.cost,
                y=logprice.scaled,
                colour=factor(set),
                group=paste(workerid, object))) +
  geom_line(stat="smooth",method = "lm", alpha = 0.1) +
  facet_wrap(~object) +
  geom_point(alpha=0.1)
```

At a glance, it looks like this might not warrant the intercept-less model we used when z-scoring within participants and objects on the full dataset.

But the model with random intercepts doesn't actually converge, so maybe it's fine to keep the model the same.

Now, in the replication design, the slopes are much steeper! Are we increasing the possibility of Type 1 errors?

### 2. Model

Given that preprocessing involves normalizing responses from each participant for each item, we would use the following model, which includes a random slope for participant but no random intercept. We include random intercepts for intensifier (the particular words). We center the continuous predictors by subtracting their means.

```{r, echo=T}
m = lmer(logprice.scaled ~ 1 + syll.centered * surprisal.centered +
    (0 + syll.centered + surprisal.centered | workerid) +
    (1 | intensifier), df)
```

### 3. Power analysis

#### Median split

```{r}
split.data = df %>% mutate(
  word.cost.contrast = ifelse(
    word.cost < median(word.cost), -1, 1),
  syll.contrast = ifelse(
    syll.centered < median(syll.centered), -1, 1),
  surprisal.contrast = ifelse(
    surprisal.centered < median(surprisal.centered), -1, 1))
m.split.combined = lmer(logprice.scaled ~ 0 + word.cost.contrast +
                          (0 + word.cost.contrast | workerid) +
                          (1 | intensifier), split.data)
m.split.syll = lmer(logprice.scaled ~ 0 + syll.contrast +
                      (0 + syll.contrast | workerid) +
                      (1 | intensifier), split.data)
m.split.surprisal = lmer(logprice.scaled ~ 0 + surprisal.contrast +
                           (0 + surprisal.contrast | workerid) +
                           (1 | intensifier), split.data)
```

We can approximate power by doing a median split on the data. This is just a rough approximation. And I'm probably implementing it wrong.

First I did a thing which is probably wrong where I assumed a single predictor that's a linear combination of surprisal and syllables. According to this (almost certainly incorrect) calculation and using [this app](https://jakewestfall.shinyapps.io/crossedpower/), we can get really, really high powe (~.99) with 72 items and 20 participants. We can get really high power (~.98) with 50 items and 50 participants. We can get high power (~.88) with 30 stimuli and 100 participants.

If syllables were our only predictor, our highest possible power with 72 items would be around .74. If we had 82 items, we could get up to .8.

If surprisal were our only predictor, our highest possible power with 72 items would be around .94. With 25 participants and 72 items we could get really high power (~.9). With 50 items and 50 participants, we could get high power (~.8).

Explanation/Summary: Once we z-score, the variance due to participants is greatly reduced, and so we need very few participants for high power. However, there is a lot of variance due to item, and we're not rescaling that. So to get high power we probably need at least 50 intensifiers. For a course feature like syllables, we need a lot more than that (~100 intensifiers).

Conclusion: It's probably fine to use 50 intensifiers and 50 participants, though we might need more intensifiers to have adequate power for the syllables factor.

Caveat: I don't really know what I'm doing. Having two continuous predictors and reducing them to a single median-split variable is kind of weird and complicated, and I'm certainly doing it wrong.

#### Bootstrapping

We could also assume that the two effects are there, and resample the data to see how often we pick up on the significance of both the predictors.

```{r}
## set up a list of data frames to draw from
workers = as.character(unique(df$workerid))
intensifiers = as.character(unique(df$intensifier))
make.indices = function(i, w) {
  return(apply(
    expand.grid(i, w), 1,
    function(v) {return(paste(v, collapse="."))}))}
pilot.data = apply(
  expand.grid(intensifiers, workers) %>%
    mutate(Var1 = as.character(Var1),
           Var2 = as.character(Var2)), 1, 
  function(v) {
    return(df %>% select(intensifier, workerid, price,
                         object, syll.centered, surprisal.centered) %>%
             filter(workerid==v[["Var2"]] & intensifier==v[["Var1"]]))})
names(pilot.data) = make.indices(intensifiers, workers)
```

```{r}
run.model = function(data) {
  preprocessed.data = data %>%
    group_by(object, workerid) %>%
    mutate(
      z = as.numeric(scale(log(price))),
      z = ifelse(is.nan(z), 0, z),
      logprice.scaled = z
    ) %>%
    select(logprice.scaled, syll.centered, surprisal.centered, workerid, intensifier, object) %>%
    ungroup
  if (F) {
    preprocessed.data %>%
      ggplot(., aes(x=(0.18*syll.centered + 0.11* surprisal.centered),
                    y=logprice.scaled,
                    group=paste(workerid, object))) +
      geom_line(stat="smooth",method = "lm", alpha = 0.1) +
      facet_wrap(~object)
  }
  resampled.model = lmer(
    logprice.scaled ~ 1 + syll.centered * surprisal.centered +
      (0 + syll.centered + surprisal.centered | workerid) +
      (1 | intensifier), preprocessed.data)
  coefficients = summary(resampled.model)$coefficients
  pvalues = c(
    syll=coefficients[["syll.centered", "Pr(>|t|)"]],
    surprisal=coefficients[["surprisal.centered", "Pr(>|t|)"]]
  )
  significance = c(
    syll = pvalues[["syll"]]<0.05,
    surprisal = pvalues[["surprisal"]]<0.05
  )
  significance["both"] = significance[["syll"]] & significance[["surprisal"]]
  return(significance)
}
bootstrap.power = function(reps=1, participantsN=28, intensifiersN=40) {
  results = sapply(1:reps, function(r) {
    resampled.workers = sample(workers, participantsN, replace=T)
    resampled.intensifiers = sample(intensifiers, intensifiersN, replace=T)
    ## for each worker, for each intensifier, what's their data?
    resampled.indices = apply(expand.grid(resampled.intensifiers, resampled.workers), 1, function(lst) {return(paste(lst, collapse="."))})
    resampled.data = do.call(rbind, pilot.data[resampled.indices])
    return(
      tryCatch(
        run.model(resampled.data),
        warning=function(e) {return(c(syllables=NA, surprisal=NA, both=NA))},
        error=function(e) {return(c(syllables=NA, surprisal=NA, both=NA))}
      )
    )
  })
  return(rowMeans(results, na.rm=T))
}
bootstrap.power(100, 28, 40)
#       syll  surprisal       both 
# 0.14141414 0.69696970 0.09090909 
# 
# bootstrap.power(100, 100, 72)
#      syll surprisal      both 
#      0.31      0.96      0.31 
# 
# bootstrap.power(100, 30, 72)
#      syll surprisal      both 
#      0.22      0.94      0.22 
# 
# bootstrap.power(100, 10, 72)
#      syll surprisal      both 
#      0.14      0.86      0.09 
```

```{r}
power.sketch.to.plot = data.frame(
  syll=c(0.14141414, 0.31, 0.22, 0.14),
  surprisal=c(0.69696970, 0.96, 0.94, 0.86),
  both=c(0.09090909, 0.31, 0.22, 0.09),
  reps=c(100, 100, 100, 100),
  participantsN=c(28, 100, 30, 10),
  intensifiersN=c(40, 72, 72, 72))
power.sketch.to.plot %>%
  gather("coefficient", "power", c(syll, surprisal, both)) %>%
  # filter(participantsN < 1000) %>%
  ggplot(., aes(x=participantsN, y=intensifiersN,
                colour=power, size=power)) +
  geom_point() +
  facet_wrap(~coefficient)
ggsave("power.sketch.png", width=10, height=4)
```

### 4. Items

See [intensifiers-list.html](./intensifiers-list.html)

### 5. Writeup

Reviewer 1 mentioned 2 important issues in our design: our choice of intensifiers and the power of our study. Our choice of intensifiers was not fully systematic or replicable, and any one of our experiments may have been underpowered (though we did replicate our findings in several pilot studies). We therefore ran a further replication with a more systematic set of intensifiers.